Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change

### 2018-09-22 LatexNotes/Notes.tex
NM
Updating latex notes. Writing up derivation of KdV so that I can calculate powers of E_0 and D_0

We will be using the truncated KdV equations (TKDV) with variable depth, in particular an abrupt depth change (ADC). The upstream depth is normalized to $1$ and the downstream (dimensionless) depth is $D_0$. Accordingly, the upstream and downstream Hamiltonians are given by
\begin{align}
\label{Hamiltonian}
& H^- = E_0^{1/2} H_3 - H_2 \\
& H^+ = D_0^{-13/4} E_0^{1/2} H_3 - D_0^{3/2} H_2 \\
\end{align}

@@

Figure \ref{Lambda16Run} shows a run with $\Lambda = 16$, with $10^7$ samples per sweep and 200 sweeps for a total of $2 \times 10^9$ nominal samples. The main parameter to vary is the upstream inverse temperature $\theta^-$. More negative $\theta^-$ gives a more significantly skewed distribution of $u$, but also results in noisier statistics due to the lower acceptance rate. I therefore chose $\theta^- = -0.2$ since it gives a good balance of the two. The top panel of the figure shows $\theta^+$ as a function of $\theta^-$ and the histograms for $\theta^- = -0.2$. The bottom panel shows microstate information, including the pdf of displacement, $u$, and the spectrum. Notice the definite skew of the downstream displacement and the decay of both upstream and downstream spectra. For this run, the upstream acceptance rate was $1\%$ and the downstream rate was $0.08\%$.

>>>>

\subsection{Nondimensionalization of the Euler system leading to KdV}

I find Johnson's derivation of KdV difficult to follow (he keeps changing his mind about what scales to use) so I am doing it myself. We start with the incompressible Euler equations in dimensional form
\begin{align}
\frac{Du}{Dt} = -\frac{1}{\rho} \pd{P}{x} \, , \qquad
\frac{Dw}{Dt} = -\frac{1}{\rho} \pd{P}{z} - g \, , \qquad
u_x + w_z = 0
\end{align}
which are supplemented with free-surface conditions
\begin{align}
&w = h_t + u h_x	\quad \text{on } z = h \\
&\phi_t + \frac{1}{2} (u^2 + w^2) + g(h-h_0) = 0  	\quad \text{on } z = h
\end{align}
and bottom conditions
\begin{equation}
w = b_t + u b_x		\quad \text{on } z = b
\end{equation}
Henceforth, we will assume the bottom topography is fixed so that the above reduces to $w = u b_x$.

We non-dimensionalize the above system by choosing a length scale of $h_0$ (the upstream depth) and a timescale of $\sqrt{h_0/g}$, which gives a velocity scale of $\sqrt{gh_0}$. 
Note: Johnson uses separate scales for horizontal versus vertical lengths and velocities, but then we rescales everything to undo this anyway. Instead, I am directly scaling all lengths and velocities on the same scales.
We write the free surface as 
\begin{equation}
h = h_0 + a \eta
\end{equation}
where $a$ is a characteristic amplitude. Thus, $\eta$ is dimensionless. We write the pressure as
\begin{equation}
P = P_a + \rho g (h_0 - z) + \rho g h_0 p
\end{equation}
so that $p$ is dimensionless. Using these scales, the dimensionless Euler equations are
\begin{align}
\frac{Du}{Dt} = -p_x \, , \qquad
\frac{Dw}{Dt} = -p_z \, , \qquad
u_x + w_z = 0
\end{align}
The free-surface conditions are
\begin{align}
&w = \eps (\eta_t + u \eta_x) \quad \text{on } z = 1 + \eps \eta \\
&\phi_t + \frac{1}{2} (u^2 + w^2) + \eps \eta = 0 \quad \text{on } z = 1 + \eps \eta
\end{align}
The bottom condition is
\begin{equation}
w = u b_x \quad \text{on } z = b
\end{equation}
Here, $\eps = a / h_0$ is a small parameter that measures the ratio of wave amplitude to depth.

\subsection{Hamiltonian structure}
We will be using the truncated KdV equations (TKDV) with variable depth, in particular an abrupt depth change (ADC). The upstream depth is normalized to $1$ and the downstream (dimensionless) depth is $D_0$. Accordingly, the upstream and downstream Hamiltonians are given by
\begin{align}
\label{Hamiltonian}
& H^- = E_0^{1/2} H_3 - H_2 \\
& H^+ = D_0^{-13/4} E_0^{1/2} H_3 - D_0^{3/2} H_2 \\
\end{align}

@@

Figure \ref{Lambda16Run} shows a run with $\Lambda = 16$, with $10^7$ samples per sweep and 200 sweeps for a total of $2 \times 10^9$ nominal samples. The main parameter to vary is the upstream inverse temperature $\theta^-$. More negative $\theta^-$ gives a more significantly skewed distribution of $u$, but also results in noisier statistics due to the lower acceptance rate. I therefore chose $\theta^- = -0.2$ since it gives a good balance of the two. The top panel of the figure shows $\theta^+$ as a function of $\theta^-$ and the histograms for $\theta^- = -0.2$. The bottom panel shows microstate information, including the pdf of displacement, $u$, and the spectrum. Notice the definite skew of the downstream displacement and the decay of both upstream and downstream spectra. For this run, the upstream acceptance rate was $0.01\%$ and the downstream rate was $0.008\%$. The total CPU time was about 33 hours.
